<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{rbhl vignette}
%\VignetteEncoding{UTF-8}
-->

```{r echo=FALSE}
library("knitr")
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(if (abs(lines[1])>1) more else NULL,
            x[lines],
            if (length(x)>lines[abs(length(lines))]) more else NULL
           )
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
})
# options(encoding = "native.enc")
knitr::opts_chunk$set(
  comment = "#>", 
  warning = FALSE, 
  message = FALSE,
  cache = FALSE
)

```

rbhl Introduction
=================

`rbhl` is an R interface to the Biodiversity Heritage Library.

## Installation

Stable version from CRAN

```{r installrbhlcran, eval=FALSE}
install.packages("rbhl")
```

Or, development version from GitHub

```{r eval=FALSE}
remotes::install_github("ropensci/rbhl")
```

Then load library

```{r loadrbhl}
library("rbhl")
```

## Output formats

You can output various formats using the `as` parameter, setting to 'json', 'xml', 'list', or 'table'. Note that not all options are available in all functions due to varying returned data objects.

Raw output, in xml format

```{r authorsearch1, messages=FALSE, warning=FALSE, eval=FALSE}
bhl_authorsearch(name='dimmock', as='xml')
```

json format

```{r authorsearch2, messages=FALSE, warning=FALSE}
bhl_authorsearch(name='dimmock', as='json')
```

Or get a list

```{r output.lines=10}
bhl_authorsearch(name='dimmock', as='list')
``` 

The default option in most functions is to parse to a data.frame

```{r authorsearch4, messages=FALSE, warning=FALSE}
bhl_authorsearch(name='dimmock')
``` 

## Some examples of function calls

### Get title metadata

```{r output.lines=10}
bhl_gettitlemetadata(1726, items = TRUE)
```

### Book search

```{r booksearch, messages=FALSE, warning=FALSE}
bhl_publicationsearch('cocos island costa rica birds')
```

### Get languages

This function gets a list of languages in which materials in BHL have been written. 

```{r}
bhl_getlanguages()
```

## Use case for BHL data

Here, we search for keywords, then plot the data using `ggplot2`

```{r cache=TRUE}
x <- bhl_publicationsearch('Selborne', year=1825, collection=4, language='eng')
pages <- bhl_pages(itemid = x$ItemID[1])
```

Examine some of the pages

```{r}
lapply(pages[7:10], "[[", "OcrText")
```

With this text data, you can do text mining of the content to answer cool questions you have. 
